---
title: "Question 1"
author: "Thamilarasan Kaveen"
date: "2023-03-27"
output: html_document
---

```{r}
# Import necessary libraries
library(ggplot2)
library(dplyr)

# Import the combined dataset
df_combined <- read.csv('D:/BSc/2nd year/Programming for Data Science/Coursework - R/combined_data.csv')

# Add leading zero to values less than 1000 in CRSDepTime column
df_combined$CRSDepTime <- ifelse(df_combined$CRSDepTime < 1000, paste0("0", df_combined$CRSDepTime),df_combined$CRSDepTime)

# Remove missing values in DepTime and ArrTime columns
df_drop <- df_combined[!is.na(df_combined$DepTime), ]
df_drop <- df_combined[!is.na(df_combined$ArrTime), ]

# Add arrival delay and departure delay to get total delay
df_drop$TotalDelay <- df_drop$ArrDelay + df_drop$DepDelay

# Add "." between the times in CRSDepTime column (ex: 0743 as 07.43)
df_drop$CRSDepTime <- gsub("^(.{2})(.*)$", "\\1.\\2", df_drop$CRSDepTime)

# Check class of CRSDepTime
class(df_drop$CRSDepTime)

# Convert it to numeric
df_drop$CRSDepTime <- as.numeric(df_drop$CRSDepTime)
```

# To find the best time of the day to fly

```{r echo=TRUE}
# Create timebins with each bin having a width of 3 hours using the CRSDepTime column
Timebins <- cut(df_drop$CRSDepTime, breaks=c(00.00, 03.00, 06.00, 09.00, 12.00, 15.00, 18.00, 21.00, 24.00))

# Merge the Timebins column with the dataset
df_drop['Timebins'] <- Timebins
head(df_drop)

# Group by Timebins and get the median delays in a table
time_avg_delay <- df_drop %>% group_by(Timebins) %>% 
  summarise(MedianDelay=median(TotalDelay),
            .groups = 'drop')
time_avg_delay
```

# Create side by side boxplots
```{r echo=TRUE, warning=FALSE}
ggplot(data=df_drop, aes(x=TotalDelay, y=Timebins)) + geom_boxplot(fill="red") +xlim(c(-100,100))+
ggtitle('Horizontal Boxplots of Total Delay by Time of day') + ylab('Timebins(Hours)')
```

# The timebin 03:00 - 06:00 is the best time of the day to minimise delays

# To find the best day of the week to fly

```{r echo=TRUE}
# Group by day of week and get the mean delays in a table
day_avg_delay <- df_drop %>% group_by(DayOfWeek) %>% 
  summarise(AverageDelay=mean(TotalDelay),
            .groups = 'drop')
day_avg_delay
```

# Create a bar plot with day on x axis and delay on y axis
```{r echo=TRUE}
ggplot(day_avg_delay, aes(x = DayOfWeek, y = AverageDelay, fill=DayOfWeek)) + geom_bar(stat = "identity") +
 scale_x_continuous(breaks = 1:7, labels=c('Mon', 'Tue', 'Wed', 'Thu', 'Fri', 'Sat', 'Sun')) + 
 theme(legend.position = "none") + ggtitle("Bar plot showing average delay for each day") + 
 xlab("Day of Week") + ylab("Average Delay(min)")
```

# Saturday is the best day to fly to minimise delays

# To find the best time of the year to fly

```{r echo=TRUE}
# Group by month and get the mean delays in a table
month_avg_delay <- df_drop %>% group_by(Month) %>% 
  summarise(AverageDelay=mean(TotalDelay), .groups = 'drop')

month_avg_delay
```

# Create a line plot with month on x axis and delay on y axis
```{r echo=TRUE}
ggplot(month_avg_delay, aes(x = Month, y = AverageDelay)) + geom_line(color = 'blue') +
  scale_x_continuous(breaks = 1:12, labels=c('Jan', 'Feb', 'Mar', 'Apr', 'May', 'Jun', 'Jul', 'Aug', 'Sep', 'Oct', 'Nov', 'Dec')) +
  ggtitle('Line plot showing average delay for each month') + ylab('Average Delay(min)')
```

# September and November are the best months of the year to fly to minimise delays