---
title: "Untitled"
author: "Thamilarasan Kaveen"
date: "2023-03-30"
output: html_document
---

```{r}
# Import necessary libraries
library(dplyr)
library(caTools)
library(ggplot2)
library(caret)
library(tidymodels)
library(ROCR) # Library needed to plot the ROC curve

# Import the combined dataset
df_combined <- read.csv('D:/BSc/2nd year/Programming for Data Science/Coursework - R/combined_data.csv')

# Check the data types
str(df_combined)

# Drop unnecessary columns
data <- subset(df_combined, select = -c(UniqueCarrier,FlightNum,TailNum,Origin,Dest,Cancelled,CancellationCode,Diverted))

# Check for null values
sapply(data, function(x) sum(is.na(x)))

# Remove null values
data <- data[complete.cases(data),]
```

# Feature selection
```{r echo=TRUE}
# Import reshape2 library
library(reshape2)

# Calculate correlation coefficients between all the variables, rounded to 2 decimal places
cor_df <- round(cor(data[sapply(data, is.numeric)]), 2)

# Melt the data frame
melted_cor <- melt(cor_df)

head(melted_cor)

# Create correlation heatmap
ggplot(data = melted_cor, aes(x=Var1, y=ordered(Var2, levels = rev(sort(unique(Var2)))), fill=value)) + geom_tile() +
  geom_text(aes(Var2, Var1, label = value), size = 2) +
  scale_fill_gradient2(low = "blue", high = "red", limit = c(-1,1), name="Correlation") + 
  ggtitle('Correlation heatmap') +  theme(axis.title.x = element_blank(),
                                          axis.title.y = element_blank(),
                                          axis.text.x= element_text(angle = 90))

```

# Splitting data
```{r echo=TRUE}
# Create new boolean column based on value in ArrDelay column. If ArrDelay > 15, 1 is assigned, else 0 is assigned
data$ArrivalDelayed <- ifelse(data$ArrDelay > 15, 1, 0)

# Check the class balance
table(data$ArrivalDelayed)
```
# As you can see the classes seem to have imbalances. Accuracy will not be considered as a metric due to this.

# Splitting data
```{r}
# Select the features with the target variable
X <- subset(data, select = c(Month, DayOfWeek, CRSDepTime, DepDelay, TaxiOut))
y <- subset(data, select = c(ArrivalDelayed))

# Split X and y into training and testing sets
set.seed(2)
sample_split <- sample.split(data$ArrivalDelayed, SplitRatio = 0.75)
X_train <- subset(X, sample_split == TRUE)
X_test <- subset(X, sample_split == FALSE)
y_train <- subset(y, sample_split == TRUE)
y_test <- subset(y, sample_split == FALSE)
```

# Scaling the dataset
```{r}
prepro_train <- preProcess(X_train, method = c('center', 'scale'))
X_train <- predict(prepro_train, X_train)
prepro_test <- preProcess(X_test, method = c('center', 'scale'))
X_test <- predict(prepro_test, X_test)
```

# Model development and prediction
```{r echo=TRUE}
# Train the model
log_reg <- glm(unlist(y_train) ~., family = binomial, data = X_train)

# Predict
y_pred <- predict(log_reg, X_test, type = "response")
y_pred_binary <- ifelse(y_pred > 0.5, 1, 0)
head(y_pred_binary)
```

# Model evaluation
```{r echo=TRUE}
# Convert y_test and y_pred as factors to plot the confusion matrix
y_test_factor <- as.factor(y_test$ArrivalDelayed)
y_pred_factor <- as.factor(y_pred_binary)

# Create a confusion matrix and a classification report
matrix_report <- caret::confusionMatrix(data = y_pred_factor, reference = y_test_factor)
matrix_report
```
# Sensitivity and Specificity are very high. Detection rate is reasonably good.
 
```{r echo=TRUE}
matrix_report_df <- data.frame(y_test_factor, y_pred_factor)

cm <- conf_mat(data = matrix_report_df, truth = y_test_factor, estimate = y_pred_factor)

autoplot(cm, type = "heatmap") + scale_fill_gradient(low = "pink", high = "cyan") +
        ggtitle('Confusion Matrix') + xlab('Actual Delay') + ylab('Predicted Delay')
```

# ROC curve
```{r echo=TRUE}
roc_pred <- prediction(y_pred, y_test$ArrivalDelayed)
roc_perf <- performance(roc_pred, 'tpr','fpr')
plot(roc_perf)
```

# AUC value
```{r echo=TRUE}
auc.perf = performance(roc_pred, measure = "auc")
print(paste0("auc = ", auc.perf@y.values))
```

# AUC is very high (0.956) therefore model is a good fit to predict arrival delays

