---
title: "Untitled"
author: "Thamilarasan Kaveen"
date: "2023-03-29"
output: html_document
---

```{r}
# Import necessary libraries
library(ggplot2)
library(dplyr)
library(tidyr)

# Import the combined dataset
df_combined <- read.csv('D:/BSc/2nd year/Programming for Data Science/Coursework - R/combined_data.csv')
```

# Date Time
```{r echo=TRUE}
# Remove rows that contain null values in arrival or dep delays
df_edit <-df_combined[!is.na(df_combined$DepDelay | df_combined$ArrDelay),]

# Add "." between the times in CRSDepTime column (ex: 0745 as 07:45)
df_edit$CRSDepTime <- gsub("^(.{2})(.*)$", "\\1:\\2", df_edit$CRSDepTime)

# Combine Year, Month, and DayofMonth column to create a Date column
df_edit$Date<-as.Date(with(df_edit,paste(Year,Month,DayofMonth,sep="-")),"%Y-%m-%d")

# Create new column called DateTimeCRSDepTime by combining Date and CRSDepTime columns
df_edit$DateTimeCRSDepTime <- paste(df_edit$Date, df_edit$CRSDepTime, sep=" ")

# Create a new column called TotalDelay
df_edit$CurrentDelay <- df_edit$ArrDelay + df_edit$DepDelay

# Remove unnecessary columns
df_edit = subset(df_edit, select = c(TailNum, CurrentDelay, Origin, Dest, DateTimeCRSDepTime))

# Sort the dataset by TailNum and then DateTimeCRSDepTime
df_edit <- df_edit[order(df_edit$TailNum, df_edit$DateTimeCRSDepTime), ]

# Create a new column which shows the previous delay by lagging the rows by 1
df_edit$PreviousDelay <- dplyr::lag(df_edit$CurrentDelay, n=1)
head(df_edit)

```

# Create a scatter plot to see if there is a relationship between delays in the previous flight and current flight

```{r echo=TRUE}
library(scattermore)
ggplot(df_edit, aes(x = PreviousDelay, y =CurrentDelay)) + geom_scattermore() +
                  ggtitle('Scatterplot showing delays in the previous flight vs delays in current flight')
```
# It is not possible to clearly say that there is a relationship since the points are all over the place.

# Create a crosstab with probabilities
```{r echo=TRUE}
# Assign 0s and 1s for TotalDelay and PreviousDelay. If Delay > 0 then 1 is assigned, else 0 is assigned
df_edit$PreviousDelay <- ifelse(df_edit$PreviousDelay > 15, 1, 0)
df_edit$CurrentDelay <- ifelse(df_edit$CurrentDelay > 15, 1, 0)
head(df_edit)

# Assign 0s and 1s for TotalDelay and PreviousDelay. If Delay > 0 then 1 is assigned, else 0 is assigned
crosstab <-table(df_edit$PreviousDelay,df_edit$CurrentDelay)

# Crosstab with probabilities of the events occuring
prop.table(crosstab, 1)
```
# 0.580 is the probability that there is a delay in the current airport given that there has been a delay in the previous airport
# 0.420 is the probability that there is no delay in the current airport given that there has been a delay in the previous airport

# 0.58 > 0.42 therefore the crosstab shows that cascading delays in the previous airport create delays in the current airport