---
title: "Question 2"
author: "Thamilarasan Kaveen"
date: "2023-03-28"
output: html_document
---

```{r echo=TRUE}
# Import necessary libraries
library(ggplot2)

# Import the combined dataset and plane dataset
df_combined <- read.csv('D:/BSc/2nd year/Programming for Data Science/Coursework - R/combined_data.csv')
plane_data <- read.csv('D:/BSC/2nd year/Programming for Data Science/Coursework - R/plane-data.csv')

# Remove missing values in DepDelay and ArrDelay columns
df_drop <- df_combined[!is.na(df_combined$DepTime), ]
df_drop <- df_combined[!is.na(df_combined$ArrTime), ]

# Add arrival delay and departure delay to get total delay
df_drop$TotalDelay <- df_drop$ArrDelay + df_drop$DepDelay

# Check datatypes of columns in both datasets
str(plane_data)
str(df_drop)
```
# We can see that the column "Tailnum" is present in both the datasets but they don't match
# Therefore the 't' and 'n' in tailnum is capitalised in the plane dataset

```{r echo=TRUE}
colnames(plane_data)[1] <- "TailNum"
str(plane_data)

# Merge the two datasets and name it as df_merge_plane
df_merge_plane <- merge(df_drop, plane_data, by="TailNum")
head(df_merge_plane)

# Remove rows that don't contain  manufacture year
df_merge_plane <- df_merge_plane[!is.na(df_merge_plane$year), ]

# We check all the unique entries
unique(df_merge_plane$year)
```
# We can see that there are rows with no particular year

```{r echo=TRUE}
# Remove rows that contain ' ', '0000' and 'None'
df_merge_plane<-df_merge_plane[!(df_merge_plane$year=="" | df_merge_plane$year=="None" | df_merge_plane$year=="0000"),]

# Group by year and get the mean delays in a table
age_avg_delay <- df_merge_plane %>% group_by(year) %>% 
  summarise(AverageDelay=mean(TotalDelay),
            .groups = 'drop')
age_avg_delay
```

# Create a line plot with year on x axis and average total delay on y axis
```{r echo=TRUE}
ggplot(age_avg_delay, aes(x = year, y = AverageDelay, group = 1)) + geom_line(color = 'red') +
         xlab('Year of manufacture of planes') + ylab('Average Delay(min)') +
         ggtitle('Line plot showing average delay against year of manufacture') + 
         theme(axis.text.x = element_text(angle = 90))
```

# Calculate the correlation coefficient between year of manufacture and average delay
```{r echo=TRUE}
# Convert year to integer
df_merge_plane$year <- as.integer(df_merge_plane$year)
cor(df_merge_plane$year, df_merge_plane$TotalDelay)
```

# There seems to be no particular trend in the line graph
# The correlation coefficient between year of manufacture and total delay is -0.002 which is very close to zero
# Hence, there is no relationship between year of manufacture and total delay
# Older planes do not suffer more delays